Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-06 paper/paper_new.tex
Updates from Overleaf
Edited the abstract and Intro.

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source particularly for researchers seeking use apply achine learning to the vast collections of digitized specimens. Unfortunately, the metadata available is often sparse and, at times, erroneous; and it is prohibitively expensive to address these limitations through traditional, manual means. This paper reports on a research applying machine-driven approaches to analyze digitized fish images and extract a variety of important features. The digitized fish specimen are being analyzed as part of Biology Guided Neural Networks (BGNN) initiative, where researchers are developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. The automatically generated metadata is crucial for identifying high-quality images for the neural network to support predictive analyses.%predictive analytics?
Research combining machine learning  and  image  informatics techniques allow us to rapidly enrich existing metadata associated for a sample of 7244 images from the Illinois Natural History Survey (INHS). The sample also includes 22 manually generated image-quality metadata properties, aiding our comparison.  Results show we can  accurately generate metadata for six key %6? 
metadata properties: fish quantity, fish orientation, image scaling based on ruler identification and measurement, and general image quality metrics. Results also show can also x, y, z.
%to confirm list of six properties.
%x, y, z = any other key results. jg will add from results section, unless someone else does first. 
The automatic process out-performs humans in terms time and labor cost, and provides a novel solution for leveraging digitized specimens for machine learning. This research has implications for harnessing the tens of thousands of digitized specimens stored in open access repositories found wold-wide.
%world wide [*across the globe] or we may not even need this part as open access is global access

@@

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images~\cite{beaman2012mass}. Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens~\cite{page2015digitization}. These digitized specimen are generally accessible through global, open access repositories supporting digital library services, such as browsing, search and retrieval, and preservation. These digitized renderings of these rich collections permit researchers, educators, students, and the general public to examine collections on a previously unattainable scale. Moreover, the digitized instantiations present a pathway for applying machine learning and making new scientific discoveries.

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images (e.g. low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements) provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata~\cite{biodiv_info_standards}, detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods %machine learning? or okay here
to examine thousands of images
to determine if, for example, a specific fish grows to different lengths
in different habitats, or the impact of temporal factors the size of a dorsal fin.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

Since the digital collections may each contain tens of thousands of images, manually producing additional image quality metadata for each digitized specimen is both labor-and time intensive and prohibitively expensive. Methods for automatically computing metadata are therefore needed to fully exploit biological image repositories for scientific discovery.
%we are both extracting and calculating. Just to double check then, our definition/understanding of computing here includes both of these, or should we focus on extraction, with the idea that calculating is the extra novel thing added, the application of math-the groovy stuff you guys added.  
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics 
techniques,
%tools and [okay to just say image informatics techniques instead of "image informatics tools and techniques" - thinking techniques implies tools.]
we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection~\cite{INHS}.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
%Dave/Joel, the section above, from "we can accurately ..." to the figures" to me is more results. I guess it's okay to give this in introduction. Just wanted to raise this point. I sometimes give a paraphrase too, but just good to see if it is a little too much or okay with you both. I think it's okay as I re-read it, but good to have someone check.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results, along with discussion, are presented. The conclusion highlights key findings and identifies next steps.

@@

\begin{table*}[t]
    \centering
    \caption{Metadata properties (* indicates higher order properties)}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale|* & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|contrast|* & Per Fish & Float & The contrast between foreground and background intensities within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis|* & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value|* & Per Fish & Integer & Fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length|* & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side|* & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table*}

>>>>

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source, particularly for researchers seeking to apply machine learning to the vast collections of digitized specimens. Unfortunately, the available metadata is often sparse and, at times, erroneous; and it is prohibitively expensive to address these limitations through traditional, manual means.
This paper reports on research that applies machine-driven approaches to
analyzing digitized fish images and extracting various of important
features from them. The digitized fish specimens are being analyzed as part of the Biology Guided Neural Networks (BGNN) initiative, which is developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. Automatically generated metadata is crucial for identifying
the high-quality images needed for the neural network's predictive
analyses.
%predictive analytics?
Methods that  combine machine learning  and  image  informatics techniques allow us to rapidly enrich the existing metadata associated with each of the
7,244 images from the Illinois Natural History Survey (INHS). The sample
images each include 22 manually generated metadata properties, which
aid our study.  Results show we can  accurately generate metadata for
twenty key %6? 
metadata properties, for example: fish quantity, fish orientation,
image scaling based on
ruler identification and measurement, and general image quality metrics
(brightness and contrast). Results also show can also x, y, z.
%to confirm list of six properties.
%x, y, z = any other key results. jg will add from results section, unless someone else does first. 
The automatic process out-performs humans in terms of time and labor cost,
as well as accuracy, and provides a novel solution for leveraging digitized
specimens for machine learning. This research has implications for harnessing the tens of thousands of digitized specimens stored in open-access repositories found wold-wide.
%world wide [*across the globe] or we may not even need this part as open access is global access

@@

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images~\cite{beaman2012mass}. Additionally, initiatives, such as the National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program,
% (which ran for nearly a decade),
have supported the digitization and curation of tens of thousands of biological specimens~\cite{page2015digitization}. These digitized specimens are generally accessible through global, open-access repositories that support digital library services, such as browsing, search and retrieval, and preservation. The digitized renderings of these rich collections permit researchers, educators, students, and the general public to examine biological specimens
on a previously unattainable scale. Moreover, the digitized instantiations present a pathway for applying machine learning and making new scientific discoveries.

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images (e.g. low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements) provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata~\cite{biodiv_info_standards}, detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods %machine learning? or okay here
to examine thousands of images
to determine if, for example, a specific fish grows to different lengths
in different habitats, or the temporal evolution of an anatomical feature,
e.g.~the size of a dorsal fin.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

Since digital collections may each contain tens of thousands of images, manually producing image-related metadata for each digitized specimen is both labor- and time-intensive and prohibitively expensive. Methods for automatically computing metadata are therefore needed to fully exploit biological image repositories for scientific discovery.
%we are both extracting and calculating. Just to double check then, our definition/understanding of computing here includes both of these, or should we focus on extraction, with the idea that calculating is the extra novel thing added, the application of math-the groovy stuff you guys added.  
As a step towards improving metadata in research specimen image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimens.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which is developing a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics 
techniques,
%tools and [okay to just say image informatics techniques instead of "image informatics tools and techniques" - thinking techniques implies tools.]
we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
have been  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection~\cite{INHS}.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
%Dave/Joel, the section above, from "we can accurately ..." to the figures" to me is more results. I guess it's okay to give this in introduction. Just wanted to raise this point. I sometimes give a paraphrase too, but just good to see if it is a little too much or okay with you both. I think it's okay as I re-read it, but good to have someone check.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results, along with discussion, are presented. The conclusion highlights key findings and identifies next steps.

@@

\begin{table*}[t]
    \centering
    \caption{Metadata properties (* indicates higher order derived properties)}
    \label{tab:properties}
    \begin{tabular}{cccp{0.5\linewidth}}
        \toprule
        \textbf{Property} & \textbf{Association} & \textbf{Type} & \textbf{Explanation}\\
        \midrule
        \verb|has_fish| & Overall Image & Boolean & Whether a fish was found in the image.\\
        \verb|fish_count| & Overall Image & Integer & The quantity of fish present.\\
        \verb|has_ruler| & Overall Image & Boolean & Whether a ruler was found in the image.\\
        \verb|ruler_bbox| & Overall Image & 4 Tuple & The bounding box of the ruler (if found).\\
        \verb|scale|* & Overall Image & Float & The scale of the image in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|bbox| & Per Fish & 4 Tuple & The top left and bottom right coordinates of the bounding box for a fish.\\
        \verb|background.mean| & Per Fish & Float & The mean intensity of the background within a given fish's bounding box.\\
        \verb|background.std| & Per Fish & Float & The standard deviation of the background within a given fish's bounding box.\\
        \verb|foreground.mean| & Per Fish & Float & The mean intensity of the foreground within a given fish's bounding box.\\
        \verb|foreground.std| & Per Fish & Float & The standard deviation of the foreground within a given fish's bounding box.\\
        \verb|contrast|* & Per Fish & Float & The contrast between foreground and background intensities within a given fish's bounding box.\\
        \verb|centroid| & Per Fish & 4 Tuple & The centroid of a given fish's bitmask.\\
        \verb|primary_axis|* & Per Fish & 2D Vector & The unit length primary axis (eigenvector) for the bitmask of a given fish.\\
        \verb|clock_value|* & Per Fish & Integer & Fish's primary axis converted into an integer ``clock value'' between 1 and 12.\\
        \verb|length|* & Per Fish & Float & The length of a fish in \(\frac{\mathrm{pixels}}{\mathrm{cm}}\).\\
        \verb|mask| & Per Fish & 2D Matrix & The bitmask of a fish in 0's and 1's.\\
        \verb|pixel_analysis_failed| & Per Fish & Boolean & Whether the pixel analysis process failed for a given fish. If \verb|true|, \verb|detectron|'s mask and bounding box were used for metadata generation.\\
        \verb|score| & Per Fish & Float & The percent confidence score output by \verb|detectron| for a given fish.\\
        \verb|has_eye| & Per Fish & Boolean & Whether an eye was found for a given fish.\\
        \verb|eye_center| & Per Fish & 2 Tuple & The centroid of a fish's eye.\\
        \verb|side|* & Per Fish & String & The side (i.e.\ \verb|'left'| or \verb|'right'|) of the fish that is facing the camera (dependent on finding its eye).\\
      \bottomrule
\end{tabular}
\end{table*}
